---
title: "datacleaning_5"
output: html_document
---

Making some changes to test pushes 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

crtl/alt/i is the keyboard shortcut for a new chunk

``` {r librarycalls, message = FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

## Data Sources

Read in the data from [KNB] https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1" 

```{r}
catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

glimpse(catch_original)
```

## Clean and Reshape Data

*remove unnecessary columns
*check column types
*reshape data

```{r}
catch_data <- catch_original %>% 
  # select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  select(-All, -notesRegCode)

glimpse(catch_data)



```

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.integer(Chinook))
```
 error, 1 chinook was coerced into n/a--code below detects where the issue is

```{r}
i <- which(is.na(catch_clean$Chinook))
i
```

Fixing it: First, replacing the I with 1, then, changing case type to integer

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.integer(Chinook))
```

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "catch")
```

```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Region, values_from = catch)
```

```{r}
#catch_long <- catch_long %>% 
  #rename(catch_thousands = catch)
  #mutate(catch_thousands = catch_thousands * 1000)
```

```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(catch_mean = mean(catch))
```

```{r}

region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

```

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))
```

```{r}
region_defs <- region_defs %>% 
  rename(Region = code, Region_Name = mgmtArea)
```


```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region"))

head(catch_joined)
```

```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAI-039"))

sites_df %>% 
  separate(site, c("island", "site_number"), "-")
```


```{r}
cities_df <- data.frame(city = c("Juneau AK", 
                                 "Sitka AK", 
                                 "Anchorage AK"))

cities_df %>% 
  separate(city, c("city", "state_code"), " ")
```

```{r}
dates_df <- data.frame(year = c("1930",
                                "1930",
                                "1930"),
                       month = c("12",
                                "12",
                                "12"),
                       day = c("14",
                               "15",
                               "16"))

dates_df %>% 
  unite(date, year, month, day, sep = "-")
```


## Join to Regions and Definitions
